Fluent statistical computing interfaces for biological data analysis

### 2020-10-26 response/response.Rmd

---
title: Thesis Amendments
author: Stuart Lee
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

First, I would like to thank my examiners for their constructive feedback and insights. I am excited to incorporate their ideas into my future research. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

## Professor Levi Waldron {-}

Thank you for your comments and identification of typographic errors.

> p. 1 Introduction: it’s -> its

Amended.

> p. 16 “Using plyranges and the the”

Amended.

> p. 34: use of `dir` for a directory path string when `dir()` is a function returning contents of a directory path could be confusing. However, this chapter is already peer-reviewed and published so I would not change it.

Noted. I can change this in the workflow for next Bioconductor release.

> P. 43-44: Why is the genome (“hg38”) set using an endomorphic function, but style (“UCSC”) set using a replacement function (`seqlevelsStyle`)?

At the time of writing, `plyranges` does not have any genome restyling functions, hence why the replacement function was used.

> p. 77: Tensorflow “non-liner embedding” p. 77-78” “a lot of” -> many

Amended.

> p. 80: “its bounded above” ->it’s

Amended.

> p. 82: “Like, when, using” -> Like when using “While our software, liminal is able “ -> While our software, liminal, is able

Amended.

>Section 5.3.1 “Finding Gestalt: focus and context” - this is probably a clever metaphor, but I didn’t understand it.

This was a reference to Buja, Cook, and Swayne (1996) paper mentioned in the paragraph above. Here "Gestalt" refers to Gestalt theory in perception and psychology, it refers to identifying patterns in visual forms.

> p. 84: “Whether, points are near, or far” -> remove both commas

Amended.

>Figure 5.2 video https://player.vimeo.com/video/43963590 does not exist. The URL given in Table 5.1 is different and correct.

Amended.

> p. 98: “We have shown in the case studies, that” -> no comma

Amended.

## Assistant Professor Stephanie C. Hicks {-}

Thank you for your thorough review. I am looking forward to incorporating your suggestions into future work.

### General Comments

### 2020-10-27 Rmd/01-intro.Rmd
modify intro

(ref:wflow-integrate) Data from multiple assays can be combined together using modelling and transformations. Data and model results can be combined, tidied and transformed to have a common representation across different granularities of the genome.

@@

Chapter \@ref(ch:plyranges) shows that the *tidy data* semantic is applicable
to range-based genomics data and develops a *fluent* interface to transforming
it called **plyranges**. The software provides a framework to an assist
an analyst to compose queries on genomics datasets. This chapter has been 
published as @Lee2019.

@@

It is rare that a biological data analysis will involve
a single measurement assay or that only one aspect of a measurement assay
will be of interest to the biological question under study (figure \@ref(fig:wflow-integrate)). While there are 
many approaches to integrating data sets from multiple assays using multivariate
statistical techniques [@Meng2016-pz], in chapter \@ref(ch:fluentGenomics)
we describe a simple end-to-end workflow for integrating 
results along the genome using **plyranges**. This workflow also shows that
our grammar based approach does not impair interoperability between
the **tidyverse** and Bioconductor approaches, and in fact they work seamlessly
together. This chapter has been published as @Lee2020-ie.

>>>>

(ref:wflow-integrate) Data from multiple assays can be combined together using modelling and transformations to gain biological insight. This is achieved either via a statistical technique or more simply via joining data and model results so they have a common representation across different granularities of the genome.

@@

Chapter \@ref(ch:plyranges) shows that the *tidy data* semantic is applicable
to in memory data measured along the genome and develops a *fluent* interface to transforming
it called **plyranges**. The software provides a framework to an assist
an analyst to compose queries on genomics datasets.  Our software is agnostic to how counts from bulk assays have been obtained. Indeed, we have used data obtained from both alignment and quantification based approaches throughout the thesis to perform useful analyses.  
This chapter has been  published as @Lee2019.

@@

It is rare that a biological data analysis will involve
a single measurement assay or that only one aspect of a measurement assay
will be of interest to the biological question under study (figure \@ref(fig:wflow-integrate)). While there are 
many approaches to integrating data sets from multiple assays using multivariate
statistical techniques [@Meng2016-pz; @Stein-OBrien2018-bk] and data structures to represent them [@Ramos2017-kn], there has been little thought given to the interoperability between these approaches and the **tidyverse**. In chapter \@ref(ch:fluentGenomics)
we describe a simple end-to-end workflow for integrating 
results along the genome using **plyranges**. This workflow shows that
our grammar based approach does not impair interoperability between
the **tidyverse** and Bioconductor approaches, and in fact they work seamlessly
together. This chapter has been published as @Lee2020-ie.
